---
output:
  word_document: default
  html_document: default
---
#Simple Linear Regression and Correlation

##Ariel Romero

```{r}
library(tidyverse)
library(GGally)

air = airquality

head(air)
```

a. The "air" dataset has data that measures the air quality in New York. The data captured ranges from May 1973 to September 1973.

b. There are 6 variables and 153 observations in the dataset. 

c. Yes, there is missing data. Primarily in the Ozone and Solar.R fields. 

d. The variable to most likely be the response variable is "Ozone"

In the "air2" column, there are 6 columns and 111 rows. 


```{r}
air2 = air %>% filter(!is.na(Ozone)) %>% filter(!is.na(Solar.R))

ggpairs(air2)
ggcorr(air2, label = TRUE)
```

a. Temperature is the variable that is most strongly correlated with the Ozone variable.

b. Day is the the variable that is the least strongly correlated with the Ozone variable. 

```{r}
ggplot(data=air2, aes(x=Temp, y=Ozone)) + geom_point()
```

Ozone and Temp have a positive correlation.

```{r}
model1 <- lm(Ozone ~ Temp, air2)

summary(model1)

```

a. Based on the R squared value,0.488, it seems that a good chunk of the data falls within the regression model, thus making it reasonable predictor. The p value associated with this value is less than .05, which means the the regression slope is significant. 

```{r}
confint(model1)
```

The range in which the slop coefficient likely falls is between 1.964787 and 2.913433

```{r}
ggplot(data=air2, aes(x=Temp, y=Ozone)) + geom_point() + geom_smooth(method="lm",se = FALSE, colour="red")
```

```{r}
predict(model1, newdata= data.frame(Temp=c(80)), interval="predict")
```

```{r}
ggplot(data = air2, aes(x=Day,y=Ozone))+geom_point()
```

It appears that there is no correlation between Day and Ozone. 

```{r}
model2 <- lm(Ozone ~ Day, air2)

summary(model2)
```

a. In this linear regression model does appear to be a good predictor for how Day affects Ozone. The R squared value is practically 0, which means that almost none of the data falls on within the model. The p value is also greater than .05 which makes the slope none significant. 

b. The range that the slop coefficient likely falls in is between -0.745321 and 0.7056539

```{r}
confint(model2)
```


```{r}
ggplot(data=air2, aes(x=Day, y=Ozone)) + geom_point() + geom_smooth(method="lm",se = FALSE, colour="red")
```

