---
output:
  word_document: default
  html_document: default
---

Task 1:

```{r}

#install.packages(tidyverse)
library(tidyverse)

diamondsdata = diamonds

diamonds

ggplot(diamondsdata, aes(x=carat, y=price)) + geom_point()

ggplot(diamondsdata, aes(x=carat, y=price, colour= cut)) + geom_point()

ggplot(diamondsdata, aes(x=carat, y=price, colour= cut)) + geom_point() + facet_wrap(~ color)

```

Task 2: There are 53,940 rows in the diamonds dataset. 

Task 3: The relationship between carat and price is positive. As the number of carats increases, the price also increases. 

Task 4: The majority of the premium cut diamonds are of higher carat and of higher price. 

Task 5: All color of diamonds are experiencing the same trend, as the carat increases so does the price. Color D seems to be the most expensive. There are some instances where just 1 carat can cost as much as $20,000. 

Task 6:
```{r}
library(readr)
InventoryData <- read_csv("InventoryData.csv")
View(InventoryData)


```

```{r}
inventoryA <- InventoryData %>% 
  filter(Supplier == "A")

inventoryA = mutate(inventoryA, OnHandRatio = `On Hand` / `Annual Demand`)

dim(inventoryA)

```


Task 7: In the inventoryA dataset, there are 3,695 rows.

Task 8: The code shown on the assignment sheet adds a column called "OnHandRatio" to the "inventoryA" dataset that is consists of dividing the "OnHand" amount by the "Annual Demand" amount. 

Task 9: 
```{r}

avg_cost <- InventoryData %>% 
  group_by(`Item SKU`) %>%
  summarize(SKUAvgCost = mean(`Cost per Unit ($)`))

view(avg_cost)
```

Task 10: Personally, I didn't find the concepts covered in MIS 503 to be very challenging. One thing I would like clarification on is when the would it be more appropriate to create a tibble instead of a data frame and vice versa.  

