---
output:
  word_document: default
  html_document: default
---
#Model Validation

##Ariel Romero

Task 1
```{r}
library(tidyverse)
library(MASS)
library(caret)

bike <- read_csv("hour.csv")

bike = bike %>% mutate(season = as_factor(as.character(season))) %>%
mutate(season = fct_recode(season,
"Spring" = "1",
"Summer" = "2",
"Fall" = "3",
"Winter" = "4")) %>% mutate(yr =as_factor(yr)) %>% 
  mutate(mnth=as_factor(mnth)) %>% mutate(hr=as_factor(hr))  %>% mutate(holiday = as_factor(as.character(holiday))) %>%
mutate(holiday = fct_recode(holiday,
"NotHoliday" = "0",
"Holiday" = "1")) %>% mutate(workingday = as_factor(as.character(workingday))) %>%
mutate(workingday = fct_recode(workingday,
"NotWorkingDay" = "0",
"WorkingDay" = "1")) %>% mutate(weathersit = as_factor(as.character(weathersit))) %>%
mutate(weathersit = fct_recode(weathersit,
"NoPrecip" = "1",
"Misty" = "2", "LightPrecip"= "3","HeavyPrecip" = "4")) %>%
   mutate(weekday = as_factor(as.character(weekday))) %>%
mutate(weekday = fct_recode(weekday,
"Sunday" = "0",
"Monday" = "1",
"Tuesday" = "2",
"Wednesday" = "3",
"Thursday" = "4",
"Friday" = "5",
"Saturday" = "6"))


set.seed(1234)
train.rows= createDataPartition(y=bike$count, p=.7,list = FALSE)

train = bike[train.rows,]
test = bike[-train.rows,]


```

Task 2: In the testing dataset, there are 5,212 rows. In the training dataset, there are 12167 rows. 

Task 3:
```{r}
mod1 <- lm(count ~ season + mnth + hr + holiday + weekday + temp + weathersit, bike)
summary(mod1)
```

The linear regression model to predict "count"" using the variables stated appearrs to be somewhat of a decent model. The adjusted R squared value is 0.6232 which in my opinion would deem acceptable.

Task 4 & 5
```{r}
predict_train <- predict(mod1, newdata = train)

head(predict_train)

predict_test <- predict(mod1, newdata = test)

head(predict_test)
```

Task: 6
```{r}
SSE = sum((test$count - predict_test)^2)

SST = sum((test$count - mean(test$count))^2)

1 - SSE/SST
```

The R square value on the testing set, 0.6285, is very close to to the training model R square value of 0.6232. The model is a appears to be a consistent predictor for the data. 

Task 7: In training/ testing splitting, the data is split into two subsets, typically 70% training and 30% testing. With k-fold cross validation, the data set is split in to k separate parts. The training process is then repeated k times. Each instance, one part is used as as the validation data and the rest is used for a training model.