---
output:
  word_document: default
  html_document: default
---
#Multiple Linear Regression and Special Issues

#Ariel Romero

```{r}
library(tidyverse)
library(GGally)
library(MASS)

bike <- read_csv("hour.csv")

bike = bike %>% mutate(season = as_factor(as.character(season))) %>%
mutate(season = fct_recode(season,
"Spring" = "1",
"Summer" = "2",
"Fall" = "3",
"Winter" = "4")) %>% mutate(yr =as_factor(yr)) %>% 
  mutate(mnth=as_factor(mnth)) %>% mutate(hr=as_factor(hr))  %>% mutate(holiday = as_factor(as.character(holiday))) %>%
mutate(holiday = fct_recode(holiday,
"NotHoliday" = "0",
"Holiday" = "1")) %>% mutate(workingday = as_factor(as.character(workingday))) %>%
mutate(workingday = fct_recode(workingday,
"NotWorkingDay" = "0",
"WorkingDay" = "1")) %>% mutate(weathersit = as_factor(as.character(weathersit))) %>%
mutate(weathersit = fct_recode(weathersit,
"NoPrecip" = "1",
"Misty" = "2", "LightPrecip"= "3","HeavyPrecip" = "4")) %>%
   mutate(weekday = as_factor(as.character(weekday))) %>%
mutate(weekday = fct_recode(weekday,
"Sunday" = "0",
"Monday" = "1",
"Tuesday" = "2",
"Wednesday" = "3",
"Thursday" = "4",
"Friday" = "5",
"Saturday" = "6"))

glimpse(bike)

ggcorr(bike, label=TRUE)

```

We convert "yr", "mnth", and "hr" in to factors because statistical models treat categorical data differently, the conversions ensure that that the modeling function treats the data correctly. 

Task 2: It appears that atemp and temp are the variables that are best correlated with count.

```{r}
ggplot(bike,aes(x=hr,y=count)) + geom_boxplot()

ggplot(bike, aes(x=season,y=count)) + geom_boxplot()

ggplot(bike, aes(x=yr,y=count)) + geom_boxplot()

ggplot(bike, aes(x=mnth,y=count)) + geom_boxplot()

ggplot(bike, aes(x=holiday,y=count)) + geom_boxplot()

ggplot(bike, aes(x=weekday,y=count)) + geom_boxplot()

ggplot(bike, aes(x=workingday,y=count)) + geom_boxplot()

ggplot(bike, aes(x=weathersit,y=count)) + geom_boxplot()
```

hr: Count is affected by hr because the time of day factors into whether people rent a bike. 

season: The season of the year affects count because the the timing of year plays a factor as to when people decide to rent bikes.

yr: I don't believe that yr plays affects count because it a certain year, cannot affect the number of bike rentals. 

mnth: mnth can have an effect on count because a month is tied to a specific time of year which can affect the amount of bike rentals.

holiday: holiday can affect count because it stands to reason that people will rent a bike if they have a day off or if there is a holiday special on rental bikes. 

weekday: weekday can have an affect on count because there can be a specific day of the week a store can have a special or rental count can increase on weekends. 

workingday: workingday can have an affect on count because the amountn of people who are available to rent a bike on a nonworking day increases. 

weathersit: weathersit affects count because it is likely that weather affects the count of rental because of the condition of the day.

```{r}
bike = dplyr::select(bike,-c(instant, dteday, registered, casual))

allmodel <- lm(count ~., bike)

summary(allmodel)

emptymodel <- lm(count ~1, bike)

summary(emptymodel)

```

```{r}
forwardmodel = stepAIC(emptymodel, direction ="forward", scope = list(upper=allmodel,lower=emptymodel), trace=TRUE)

summary(forwardmodel)

```

The variables that are included in the forward model are:
hr
temp
atemp   
hum         
mnth       
season     
yr          
weathersit  
windspeed   
holiday    
workingday
weekday

The adjusted R-squared value is 0.6854 and the p-value less than 0.05 which means the variables are significant. I believe that the model is acceptable in predicting count given all the variables. Many of the varibles that I thought made sense in affecting count appeared in the model. 

```{r}
backwardsmodel = stepAIC(allmodel,direction="backward", trace=TRUE)
summary(backwardsmodel)
```

The backwards model is similar in the variables, coefficients of the variables, adjusted R-squared, and in p-value. 

workingday could be directly correllated to the weekday which would hinder it from appearing in the model.

```{r}

bike = bike %>% mutate(yr= as.integer(yr)-1)

allmodel <- lm(count ~., bike)

emptymodel <- lm(count ~1, bike)

forwardmodel = stepAIC(emptymodel, direction ="forward", scope = list(upper=allmodel,lower=emptymodel), trace=TRUE)

summary(forwardmodel)

```

It does not appear as though the model has changed by converting yr as an integer.