---
output:
  word_document: default
  html_document: default
---
# Missing Data

## Ariel Romero

```{r}
options(tidyverse.quiet = TRUE)
library(tidyverse)
library(VIM)
library(mice)

grades <- read_csv("class-grades.csv")

summary(grades)
                                   
                                

```

Task 1:

Tutorial- 1 data point missing, Midterm- 3 data points missing, TakeHome- 3 data points missing, Final- 4 data points missing.

Task 2: 

```{r}
vim_plot = aggr(grades, numbers = TRUE, prop = c(TRUE, FALSE),cex.axis=.7)
```

Based on the vim plot above, it appears students are missing multiple pieces of data in the Midterm and TakeHome variables. 

Task 3:

```{r}
gradesRowsWise = grades %>% drop_na()

```

After using row-wise deletion, there are 89 rows remaining in the data frame. 
Task 4:

```{r}
gradesColWise = grades %>% select(c("Prefix","Assignment"))
```

After column-wise deletion, only 2 variables remain. 

Task5: 

For this dataset, the rows-wise deletion would be the better approach in handling the missing data. Although 10 observations are eliminated, all the variables still remain which would make for a more accurate model. 

task 6: 

```{r}
grades_imp = mice(grades, m=1, method = "pmm", seed = 12345)
#in line above: m=1 -> runs one imputation, seed sets the random number seed to get repeatable results
summary(grades_imp)
densityplot(grades_imp)
#red imputed, blue original, only shows density plots when more than 1 value the variable was imputed
#note that the density plots are fairly uninteresting given the small amount of missing data
grades_complete = complete(grades_imp)
summary(grades_complete)
```

Task 7: 

Missing data can have a signfiicant effect on the conclusion that can be made from the data. Missing data can reduce the effectiveness the analysis being conduted, it can cause bias, and ultimately make the analysis more complicated by having to take more steps to clean the data. 


